---
title: A way to simplify package use on a R project
author: Vitor Chagas
date: '2013-08-19'
slug: a-way-to-simplify-package-use-on-a-r-project
categories:
  - R
tags: []
description: ''
image: ''
---

# Introduction

One of the steps that I always do when I begin an *R* script is to check if all the needed packages are installed, and if not install them.

That way I know that the script won't crash on another computer because of a missing package (one less reason to crash :D).

Lately I've been thinking about the subject of code reuse and decided to transform my block of code in to an *R* function.

##The Code

```{r, eval = FALSE}
#««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««
#  script : loadlib.R
#  author : Vitor Chagas(VC)                last updated : 2013.08.19
#««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««««

# install & load necessary packages ----
#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<

loadlibs <- function(...) {

  necessary <- as.vector(as.character(match.call()[-1]))

  installed <- necessary %in% .packages(all.available = TRUE)

  # install
  if (length(necessary[!installed]) >=1)
    install.packages(necessary[!installed])

  # update
  update.packages()

  # load
  for(pkg in necessary)
    library(pkg, character.only=TRUE)  

  rm(necessary, installed, pkg)
}
```

## Final comments

Now I just save this code into a lib folder, and then use the following code to source every code file in there

```{r}
sapply(list.files(pattern="*.R", path="lib", full.names=TRUE),
       source,.GlobalEnv)
```
